In this paper, we have presented HiFiNet, a hierarchical fault identification network for Wireless Sensor Networks that synergistically combines edge-based classification with graph aggregation. By first analyzing temporal data at the individual node level and then refining the diagnosis through a graph attention mechanism that incorporates spatial context from neighboring nodes, HiFiNet achieves a new level of accuracy and robustness in fault detection. Our comprehensive experiments, conducted on realistic datasets with injected faults, have demonstrated the superiority of HiFiNet over methods such as DBN, LSTM-AE, and SVM. The results show a significant improvement in key performance metrics, including accuracy, F1-score, and precision, across various fault types and rates. Notably, HiFiNet exhibits greater stability as noise and data quality issues increase, a critical advantage in real-world WSN deployments. A tunable mechanism unique to HiFiNet was also explored which demonstrated the model ability to balance accuracy and power consumption.

Future work could explore the adaptation of HiFiNet to dynamic network topologies and the inclusion of a wider range of fault types. Additionally, the implementation of the proposed model on actual WSN hardware would provide further insights into its real-world performance and energy consumption. Overall, HiFiNet represents a significant advancement in the field of fault diagnosis for WSNs, offering a powerful and adaptable solution for ensuring data integrity and reliability in a wide array of applications.
